Improved Bayesian Ethnorace Prediction

### 2021-04-26 paper.Rmd
replication plots added

```{r libraries}
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(here)
library(grid)
```

```{r data}
load(here("data", "tests.rda"))
```

@@

```{=tex}
\begin{singlespace}
Quantitative social science research on race and ethnicity can be constrained by the lack of available individual-level data with these markers. In response to this issue, methods have been developed to predict individuals' race and ethnicity using Bayes' rule. Racial distributions from the US Census Surname List are combined with distributions from geolocations, such as Census blocks or zip codes, to yield predicted probabilities of individuals' race and ethnicity. I expand upon existing methods by incorporating information from a nationwide list of first names as well as from residence characteristics. Along with some other adjustments, these improvements lead to substantial gains in predictive performance when validated against official state voter files. The largest of these predictive gains are found for African Americans---a group which existing methods find difficult to predict accurately.
\end{singlespace}
```
# Introduction

@@

To overcome some of these challenges, I describe a method for imputing individual ethnoraces directly. This method uses Bayes' rule to make predictions by combining information from nationwide distributions of six ethnorace categories over other characteristics, such as names, geographies, political party identification, and others. My implementation builds off existing methods which use a similar prediction algorithm [@elliott2009; @imai2016; @voicu2018]. The method described by Imai and Khanna (2016) in particular, and the associated `R` package `wru`, has become popular in recent studies on race and ethnicity. Its application has been used in work on racial protests and voting patterns [@enos2019], disparities in campaign financing [@grumbach2020; @grumbach2020a], the impact of electoral institutions on local representation [@abott2020], and public health issues such as suicide rates [@studdert2020].

My ethnorace prediction method improves upon Imai and Khanna (2016) in several ways. Whereas their method only takes as inputs distributions over surnames, geolocation, political party, age, and gender, my implementation adds information from a nationwide list of first names [@tzioumis2018] as well as address characteristics. Additionally, I incorporate insights from the machine learning literature to further improve predictive performance. The result of these modifications is a substantial increase in predictive power compared to Imai and Khanna (2016) in validation tests. These predictive gains are particularly strong in regards to correctly classifying African American and Hispanic individuals---especially in contexts where fine-grain geolocation data are unavailable. My method is available in an easy-to-use `R` package `bper`.[^1]

@@

@@

@@

## References

>>>>

```{r libraries}
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(here)
library(grid)
library(caret)
```

```{r data}
load(here("data", "tests.rda"))
load(here("data", "cal_data.rda"))
load(here("data", "sahn_grumbach_figs.rda"))
```

@@

```{=tex}
\begin{singlespace}
Quantitative social science research on race and ethnicity can be constrained by the lack of available individual-level data with these markers. In response to this issue, methods have been developed to predict individuals' race and ethnicity using Bayes' rule. I expand upon existing methods by incorporating information into the algorithm. Along with some other adjustments, these improvements lead to substantial gains in predictive performance when validated against official state voter files. The largest of these predictive gains are found for African Americans---a group which existing methods find difficult to predict accurately. Furthmore, I apply my algorithm to previous reseach using the old methods and demonstrate the substantive implications of switching to more accurate predictions.
\end{singlespace}
```
# Introduction

@@

To overcome some of these challenges, I describe a method for imputing individual ethnoraces directly. This method uses Bayes' rule to make predictions by combining information from nationwide distributions of six ethnorace categories over other characteristics, such as names, geographies, political party identification, and others. Intuitively, an individual with a name that is highly unique to a particular ethnorace, and who lives in an area with many other people of the same ethnorace, will be given a high predicted probability of belonging to that group.

My implementation builds off existing methods which use a similar prediction algorithm [@elliott2009; @imai2016; @voicu2018]. The method described by Imai and Khanna (2016) in particular, and the associated `R` package `wru`, has become popular in recent studies on race and ethnicity. Its application has been used in work on racial protests and voting patterns [@enos2019]; disparities in campaign financing [@grumbach2020; @grumbach2020a], evictions [@hepburn2020], and voter turnout [@fraga2018]; the impact of electoral institutions on local representation [@abott2020]; and public health issues such as suicide rates [@studdert2020].

My ethnorace prediction method improves upon Imai and Khanna (2016) in several ways. Whereas their method only takes as inputs distributions over surnames, geolocation, political party, age, and gender, my implementation adds information from a nationwide list of first names [@tzioumis2018] as well as address characteristics. Additionally, I incorporate insights from the machine learning literature to further improve predictive performance. The result of these modifications is a substantial increase in predictive power compared to Imai and Khanna (2016) in validation tests. These predictive gains are particularly strong in regards to correctly classifying African Americans individuals---especially in contexts where fine-grain geolocation data are unavailable. My method is available in an easy-to-use `R` package `bper`.[^1]

@@

```{r, fig.cap="Probability Calibration Plots by Ethnorace"}
cal_api <-
  ggplot(cal_data[[12]]) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  theme_minimal() +
  labs(y = "Observed Asian Percentage",
       x = "Predicted Probability Asian",
       title = "Asian") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_black <-
  ggplot(cal_data[[11]]) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  theme_minimal() +
  labs(y = "Observed Black Percentage",
       x = "Predicted Probability Black",
       title = "Black") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_hispanic <-
  ggplot(cal_data[[14]]) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  theme_minimal() +
  labs(y = "Observed Hispanic Percentage",
       x = "Predicted Probability Hispanic",
       title = "Hispanic") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_white <-
  ggplot(cal_data[[15]]) +
  geom_line(color = "darkcyan") +
  geom_point(color = "darkcyan") +
  theme_minimal() +
  labs(y = "Observed White Percentage",
       x = "Predicted Probability White",
       title = "White") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

grid.arrange(cal_api, cal_black, cal_hispanic, cal_white,
             nrow = 2, ncol = 2)
```

@@

# Replication

```{r, fig.cap="Ethnoracial Composition of the Contributor Class (1980 - 2014)"}
sahn_grumbach_figs[[2]] %>%
  ggplot() +
  geom_col(aes(x = year, y = share, fill = pred_race)) +
  theme_minimal() +
  labs(x = 'Year', y = 'Prop. of Contributions') +
  scale_fill_grey(labels = c("Asian", "Black", "Hispanic")) +
  theme(legend.title = element_blank(), legend.position = "bottom") +
  theme(text = element_text(family = "serif")) +
  facet_wrap(~ method)
```

```{r, fig.cap="Ethnoracial Composition of the Contributor Class Versus Electorate"}
sahn_grumbach_figs[[3]] %>%
  ggplot(aes(x = race, fill = race, y = share)) +
  geom_bar(stat = "identity") +
  scale_fill_grey() +
  scale_y_continuous(breaks = c(0, .05, .1)) +
  guides(fill = F) +
  labs(x = "", y = "Proportion") +
  theme_minimal() +
  facet_wrap( ~ type, nrow = 1)  +
  theme(text = element_text(family = "serif"))
```

```{r, fig.cap="Average Total Contributions by Ethnorace"}
sahn_grumbach_figs[[4]] %>%
  mutate(r_pred_race = case_when(r_pred_race == "api" ~ "Asian",
                                 r_pred_race == "black" ~ "Black",
                                 r_pred_race == "hispanic" ~ "Hispanic",
                                 r_pred_race == "white" ~ "White"),
         pred_race = case_when(pred_race == "api" ~ "Asian",
                               pred_race == "black" ~ "Black",
                               pred_race == "hispanic" ~ "Hispanic",
                               pred_race == "white" ~ "White")) %>% 
  ggplot(aes(x = r_pred_race, y = avg_amount, fill = method)) +
  geom_bar(stat = "identity") +
  scale_fill_grey() +
  labs(x = "Contributor Ethnorace",
       y = "Average Contributions (Total $)",
       fill = "Method") +
  theme_minimal() +
  facet_grid(method ~ pred_race)  +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(angle = 35))
```

# Conclusion

@@

# References
